Demes: a specification for interchanging complex demography

### 2021-06-18 paper.tex
Discussion text.

\begin{code}
\begin{tcolorbox}
\inputminted[linenos,numbersep=5pt]{yaml}{models/IM.yaml}
\end{tcolorbox}
\captionof{listing}{An IM model.}
\label{code:im_model}
\end{code}

@@

@@

This is an unstructured list of things we should mention:

    \item
    % JK: I thought there was another simulation frontend somewhere..
    Previous efforts such as \texttt{coala}~\citep{staab2016coala} have focused on
    abstracting the differences between individual simulators by providing
    a frontend.

    \item
    % https://github.com/popsim-consortium/demes-paper/issues/11
    We should mention coasim~\citep{mailund2005coasim} because its model
    seems quite close to what we have. It's not suitable as a standard though
    because it uses a programming language.

>>>>

\begin{figure}
\begin{minipage}{0.48\textwidth}
\begin{tcolorbox}
\inputminted[fontsize=\scriptsize,linenos,numbersep=5pt]{yaml}{models/IM.yaml}
\end{tcolorbox}
\begin{tcolorbox}
\includegraphics[width=\textwidth]{fig/IM}
\end{tcolorbox}
\end{minipage}\hfill
\begin{minipage}{0.48\textwidth}
\begin{tcolorbox}
\inputminted[fontsize=\scriptsize,linenos,numbersep=5pt]{yaml}{models/IM-resolved.yaml}
\end{tcolorbox}
\end{minipage}\\
\captionof{listing}{An IM model.}
\label{code:im_model}
\end{figure}

@@

\subsection*{Design rationale}

This is an unstructured list of things we should mention in terms
of why certain decisions were made

@@

The difficulty of describing complex demographic models
for population genetic simulations has long been acknowledged
~\citep[][e.g.]{antao2007modeler4simcoal2}, % jk - any older refs for this?
and a number of different methods have been proposed to
make such simulations more accessible and less error prone.
Several Graphical User Interface (GUI) and visualisation
methods have been developed, which greatly facilitate
interpretation~\citep{mailund2005coasim,antao2007modeler4simcoal2,
ewing2010msms,zhou2018popdemog}. However, these methods
currently have little traction as they are all either directly coupled
to an internal simulation method or to the specific syntax
of a given simulator. Another approach that has been taken to
simplify running such simulations is to provide an API in a
scripting language such as Python or R, providing a flexible
interface either to a built-in simulation
engine~\citep{thornton2014cpp,kelleher2016efficient,haller2017flexible}
% jk: are/were there other simulation frontends out there?
or as a frontend for other simulators~\citep{staab2016coala}.
While APIs have many advantages and have grown in popularity in
recent years, the demographic model descriptions are tied
to the details of both the underlying simulation engines
and the programming language in question.

Stable and healthy software ecosystems require standard interchange
formats, allowing for the development of high-quality and long-lasting
tools that produce and consume the standard.
Demographic models are a key part of population genetics research,
and to date the transfer of inferred models to downstream simulations
has been ad-hoc, and conversions between the many different ways
of expressing such models both labour intensive and fraught with errors.
The proposed Demes standard is an attempt to bridge this gap
between inference and simulation, and also to provide the foundations
for a sustainable ecosystem of tools built around this data model.
Table~\ref{tab-software} shows some initial infrastructure that we have
built as part of developing Demes, but many other useful tools
can be envisaged that either produce or consume this format.

% Could finish up here by talking about reproducibility?
